---
title: "project_ordinalregression"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr) 
require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)
library(datasets)
library(caret)
library(pROC)

library(AER)    # applied econometrics with R
library(plm)    # panel-lm
library(stargazer)      # popular package for regression table-making
library(lattice)

library(GGally)
library(standardize)
library(QuantPsyc)

library(tidyverse)
library(factoextra)

library("lmtest")
library("sandwich")
library("nnet")

library(Matrix)
library(MASS)
library(brant)
```


## Load data


```{r}
#read games dataset 
#games_stat = read.csv(file = 'games_stats_150000.csv', header= TRUE)
games_stat = read.csv(file = 'games_stats_analysis.csv', header= TRUE)
head(games_stat)
```


```{r}
# Column names of dataset
colnames(games_stat)
```
```{r}

games_stat_t <- games_stat %>%
  rename(owner_cat =total_owner_cat, user_age=required_age, lang = languages_supported, cat_count = categories_count, action = genre_action, anime = genre_animation___modeling, game_dev= genre_game_development ,multiP=  genre_massively_multiplayer, sexual = genre_sexual_content,sports= genre_sports, violent =genre_violent, mac=Mac_platform,win=Window_platform,race =  genre_racing,  free=genre_free_to_play, adventure=genre_adventure, linux=Linux_platform, design =genre_design___illustration,casual =  genre_casual, indie=genre_indie)



game_regression <-
 games_stat_t %>% dplyr::select(owner_cat,lang, anime, design,game_dev, multiP, sexual, sports, violent, action, race, adventure, casual,free,mac,win,  linux,cat_count, genre_count, user_age, age, custom_user_score, is_free, initialprice, controller_support,indie)
```

```{r}
str(game_regression)
```



```{r}
game_regression$owner_cat  <- as.ordered(game_regression$owner_cat ) 

str(game_regression)

```

```{r}
summary(game_regression$owner_cat)
```


```{r}
games_regression_model <- game_regression %>% dplyr::select(c(anime,design,game_dev,sexual,sports,violent,race,adventure,win,initialprice,controller_support,owner_cat))
```


```{r}
## fit ordered logit model and store results




model <- polr(owner_cat ~ ., data = games_regression_model, Hess=TRUE)


## view a summary of the model
summary(model)
```



```{r}
pred_data<- predict(model,games_regression_model)
```


```{r}
(tab_full <- table(pred_data, games_regression_model$owner_cat))
```

```{r}
1 - sum(diag(tab_full))/sum(tab_full)
```



```{r}
(ctable <- coef(summary(model)))
```
```{r}
## calculate and store p values
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

## combined table
(ctable <- cbind(ctable, "p value" = p))
```

# Ordinal Logistic Regression Assumptions

Since the Ordinal Logistic Regression model has been fitted, now we need to check the assumptions to ensure that it is a valid model. The assumptions of the Ordinal Logistic Regression are as follow and should be tested in order:

* The dependent variable are ordered.
  We have selected number of owners as our dependent variable. It has been categorized into 3 levels namely 0 (10000 to 75000), 1(150000 to 750000) and 2(1500000 to 150000000) with 0<1<2.

* One or more of the independent variables are either continuous, categorical or ordinal.
  We have selected following varaibles as our independednt variable. 
  
* No multi-collinearity.



* Proportional odds
```{r}

brant::brant(model)
```


# Test

### Split dataset to train and test

```{r}
ind <- sample(2, nrow(games_regression_model), replace=TRUE, prob=c(0.8,0.2))
train <- games_regression_model[ind==1,]
test <- games_regression_model[ind==2,]
```
### Train model with train data

```{r}

#train$owner_cat <- as.integer(train$owner_cat)

model_train <- polr(owner_cat ~ ., data = train, Hess=TRUE)

```

```{r}
pred_train<- predict(model_train,train)
```

```{r}
(tab_train <- table(pred_train, train$owner_cat))
```
```{r}
1 - sum(diag(tab_train))/sum(tab_train)
```


### Predict

```{r}
pred<- predict(model_train,test)
```

### Confusion Metrics

```{r}
(tab_test <- table(pred, test$owner_cat))
```

### Misclassification Error

```{r}
1 - sum(diag(tab_test))/sum(tab_test)
```

### Metrics

```{r}
cm <- as.matrix(confusionMatrix(pred, test$owner_cat))
n = sum(cm) # number of instances
nc = nrow(cm) # number of classes
rowsums = apply(cm, 1, sum) # number of instances per class
colsums = apply(cm, 2, sum) # number of predictions per class
diag = diag(cm)  # number of correctly classified instances per class 

precision = diag / colsums 
recall = diag / rowsums 
f1 = 2 * precision * recall / (precision + recall)
```

```{r}
print(" ************ Confusion Matrix ************")
print(cm)
print(" ************ Diag ************")
print(diag)
print(" ************ Precision/Recall/F1 ************")
print(data.frame(precision, recall, f1))

```

